A statistical framework for whole-genome bisulfite-sequencing data

### 2015-05-01 Rmarkdown/wgbs_statistical_framework.Rmd
Add brief captions

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/bs-seq_protocol.pdf}
\caption{A schematic illustrating several sources of variation in a bisulfite-sequencing experiment. Each sample starts as a population of cells, with potentially different methylation patterns. Each DNA fragment is coloured by its originating cell (although this is unknown in practice). Illustrated are PCR amplication bias (unequal representation of DNA from each cell following PCR amplification), sampling variation (not all DNA fragments are sequenced), alignment errors (not all sequenced fragments are mapped or may be mapped to the wrong location, as is the case for the green fragment), and filtering during methylation calling (not reads are used, the blue read, and reads may have read-positions removed, the black reads).}
\label{fig:bs-seq_protocol}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:EPISCOPE_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Lister_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Seisenberger_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Ziller data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Ziller_plotBetaFreqPoly}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:EPISCOPE_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Lister_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Seisenberger_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Ziller_merged_plotBetaFreqPoly_CGI}
\end{figure}

>>>>

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/bs-seq_protocol.pdf}
\caption[Source of variation in bisulfite-sequencing experiments]{A schematic illustrating several sources of variation in a bisulfite-sequencing experiment. Each sample starts as a population of cells, with potentially different methylation patterns. Each DNA fragment is coloured by its originating cell (although this is unknown in practice). Illustrated are PCR amplication bias (unequal representation of DNA from each cell following PCR amplification), sampling variation (not all DNA fragments are sequenced), alignment errors (not all sequenced fragments are mapped or may be mapped to the wrong location, as is the case for the green fragment), and filtering during methylation calling (not reads are used, the blue read, and reads may have read-positions removed, the black reads).}
\label{fig:bs-seq_protocol}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity.pdf}
\caption[\emph{EPISCOPE} $\beta$-value kernel density]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity.pdf}
\caption[\emph{Lister} $\beta$-value kernel density]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity.pdf}
\caption[\emph{Seisenberger} $\beta$-value kernel density]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity.pdf}
\caption[\emph{Ziller} $\beta$-value kernel density]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data. Densities are normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly.pdf}
\caption[\emph{EPISCOPE} $\beta$-value frequency polygon]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:EPISCOPE_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly.pdf}
\caption[\emph{Lister} $\beta$-value frequency polygon]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Lister_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly.pdf}
\caption[\emph{Seisenberger} $\beta$-value frequency polygon]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Seisenberger_plotBetaFreqPoly}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly.pdf}
\caption[\emph{Ziller} $\beta$-value frequency polygon]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Ziller data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10 \times$ sequencing coverage are included.}
\label{fig:Ziller_plotBetaFreqPoly}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity_CGI.pdf}
\caption[\emph{EPISCOPE} $\beta$-value kernel density stratified by CpG islands]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity_CGI.pdf}
\caption[\emph{Lister} $\beta$-value kernel density stratified by CpG islands]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity_CGI.pdf}
\caption[\emph{Seisenberger} $\beta$-value kernel density stratified by CpG islands]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity_CGI.pdf}
\caption[\emph{Ziller} $\beta$-value kernel density stratified by CpG islands]{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10 \times$ sequencing coverage are included. `Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly_CGI.pdf}
\caption[\emph{EPISCOPE} $\beta$-value frequency polygon stratified by CpG islands]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{EPISCOPE} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:EPISCOPE_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly_CGI.pdf}
\caption[\emph{Lister} $\beta$-value frequency polygon stratified by CpG islands]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Lister} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Lister_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly_CGI.pdf}
\caption[\emph{Seisenberger} $\beta$-value frequency polygon stratified by CpG islands]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Seisenberger} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Seisenberger_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly_CGI.pdf}
\caption[\emph{Ziller} $\beta$-value frequency polygon stratified by CpG islands]{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the \emph{Ziller} data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10 \times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Ziller_merged_plotBetaFreqPoly_CGI}
\end{figure}
